\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2017-01-31 mSystems/rev2/mSys_MandM_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	new file:   figs/HostLifeStyle_StoolA_day72to122_abroad_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day0to70_before_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day123to256_returned_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day257to364_after_TAILLESS_Rank.pdf
	modified:   figs/_ed_Fig1.pdf
	modified:   figs/_ed_supfig

%%
\subsection*{Rank Stability Index}\label{sec:RSI}

@@

Finally, under the rank matrix of the aforementioned Figures there are plots relevant to the variability of the rank over time. On the one hand, the RV (Rank Variability) for a sampling point shows the absolute difference between every taxon?s rank and its accumulated abundance rank (the overall rank), averaged for all the taxa shown. On the other hand, the DV (Differences Variability) for a sampling point shows the absolute difference between every taxon?s rank at that time and the value that it had at the previous sampling point, averaged for all the taxa shown.

>>>>

%%
\subsection*{Rank stability and variability}\label{sec:RSI}

@@

Finally, under the rank matrix of the aforementioned Figures there are plots relevant to the variability of the rank over time. On the one hand, the RV (Rank Variability) for a sampling point shows the absolute difference between every taxon's rank and its accumulated abundance rank (the overall rank), averaged for all the taxa shown. On the other hand, the DV (Differences Variability) for a sampling point shows the absolute difference between every taxon's rank at that time and the value that it had at the previous sampling point, averaged for all the taxa shown.

### 2017-01-31 mSystems/rev2/mSys_intro_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	new file:   figs/HostLifeStyle_StoolA_day72to122_abroad_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day0to70_before_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day123to256_returned_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day257to364_after_TAILLESS_Rank.pdf
	modified:   figs/_ed_Fig1.pdf
	modified:   figs/_ed_supfig

The desire to understand the factors that influence human health and cause disease has always been one of the major driving forces of biological research. As evidence of the new "holobiont"  and "hologenome" concepts is increasing each day \cite{holo1, holo2}, research not only focuses on the human physiology but also on the microbial population that surrounds us. However, these concepts are still under debate \cite{holo3}. The human microbiome is intimately linked to our physiology with the metabolism of bile acids \cite{bileacids}, of choline \cite{choline} and key metabolites like short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}. Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some afflictions like obesity \cite{ob1, ob2} and malnutrition \cite{nutr} as well as many other diseases \cite{Moya_trends}. Recent studies have revealed that microbes also influence brain function and behaviour and are related to neurological disorders like Alzheimer's disease through the gut-brain axis\cite{mind,AD}. Recently, even a subtle condition like chronic fatigue syndrome, which has often been suggested to be a psychosomatic disease, has been associated with reduced diversity and altered composition of the gut microbiome\cite{CFS}.

>>>>

The desire to understand the factors that influence human health and cause disease has always been one of the major driving forces of biological research. As evidence of the new "holobiont"  and "hologenome" concepts is increasing each day \cite{holo1, holo2}, research not only focuses on the human physiology but also on the microbial population that surrounds us. However, these concepts are still under debate \cite{holo3}. The human microbiome is intimately linked to our physiology with the metabolism of bile acids \cite{bileacids}, of choline \cite{choline} and key metabolites like short-chain fatty acids \cite{scfa1, scfa2} which are also involved in immune system maturation \cite{scfa3, scfa4}. Human microbiota has been suggested to be closely related to diseases like type 2 diabetes \cite{diabetes2}, cardiovascular disease (CVD) \cite{CVD}, irritable bowel syndrome \cite{IBS}, Crohn's disease \cite{CD}, some afflictions like obesity \cite{ob1, ob2} and malnutrition \cite{nutr} as well as many other diseases \cite{Moya_trends}. Recent studies have revealed that microbes also influence brain function and behaviour and are related to neurological disorders like Alzheimer's disease through the gut-brain axis\cite{mind,AD}. Even a difficult to diagnose condition like chronic fatigue syndrome, which in the past had often been suggested to be a psychosomatic disease, has recently been associated with reduced diversity and altered composition of the gut microbiome\cite{CFS}.

### 2017-01-31 mSystems/rev2/mSys_results_rev2.tex
mSys: several figs&text changes which lead to REV2 DRAFT delta
On branch master changes to be committed:
	modified:   Response to Reviewer Comments.docx
	modified:   figs/Fig1.eps
	modified:   figs/Fig4.eps
	modified:   figs/Fig5.eps
	new file:   figs/HostLifeStyle_StoolA_day72to122_abroad_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day0to70_before_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day123to256_returned_TAILLESS_Rank.pdf
	new file:   figs/HostLifeStyle_h_StoolA_day257to364_after_TAILLESS_Rank.pdf
	modified:   figs/_ed_Fig1.pdf
	modified:   figs/_ed_supfig

\begin{figure}
	\centering
	\vspace*{-15mm} % Corrects overbox of the figures
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{figs/Fig1.eps}
	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored over time. The fit is shown for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}. Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude. $V$ corresponds to the y-intercept and $\beta$ to the slope of the line. The error bars (\emph{mean-axis}) are the SEM.}
	\label{fig:main1}
\end{figure}

@@

\begin{figure}
	\centering
	%\vspace*{-10mm} % Corrects overbox of the figures
	\includegraphics[width=1.0\textwidth]{figs/Fig4.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for a special period (days 72 to 122, travelling abroad) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{fig:corrank_HLS_abroad}
\end{figure}

\begin{figure}
	\centering
	%\vspace*{-10mm} % Corrects overbox of the figures \hspace*{-4.5mm}
	\includegraphics[width=1.0\textwidth]{figs/Fig5.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 123 to 256, after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{fig:corrank_HLS_returned}
\end{figure}

@@

\begin{supfig}
	\centering
	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_before.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 0 to 70, before the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{supfig:corrank_HLS_before}
\end{supfig}

\begin{supfig}
	\centering
	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_after.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated Rank Stability Index (as shown in Material and Methods) for an ordinary period (days 257 to 364, further after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{supfig:corrank_HLS_after}
\end{supfig}

>>>>

\begin{figure}
	\centering
	%\vspace*{-15mm} % Corrects overbox of the figures
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.7\textwidth]{figs/Fig1.eps}
	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored over time. The fit is shown for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}. Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude. $V$ corresponds to the y-intercept and $\beta$ to the slope of the line. The error bars (\emph{mean-axis}) are the SEM.}
	\label{fig:main1}
\end{figure}

@@

\begin{figure}
	\centering
	%\vspace*{-10mm} % Corrects overbox of the figures
	\includegraphics[width=1.0\textwidth]{figs/Fig4.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for a special period (days 72 to 122, travelling abroad) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{fig:corrank_HLS_abroad}
\end{figure}

\begin{figure}
	\centering
	%\vspace*{-10mm} % Corrects overbox of the figures \hspace*{-4.5mm}
	\includegraphics[width=1.0\textwidth]{figs/Fig5.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 123 to 256, after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{fig:corrank_HLS_returned}
\end{figure}

@@

\begin{supfig}
	\centering
	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_before.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 0 to 70, before the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{supfig:corrank_HLS_before}
\end{supfig}

\begin{supfig}
	\centering
	\includegraphics[width=1.0\textwidth]{figs/supfig_corrank_HLS_StoolA_after.eps}
	\caption{Rank variation over time for the 50 most dominant elements (taxa) and their calculated RSI (Rank Stability Index), Rank Variability (RV) and Differences Variability (DV), as detailed in Rank stability and variability in Material and Methods, for an ordinary period (days 257 to 364, further after the trip) belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife}.}
	\label{supfig:corrank_HLS_after}
\end{supfig}
